A Software Engineering Perspective on \\ OpenMP and PThreads

### 2017-08-01 sections/04-research-design.tex
Improve responses section.

\todo[inline,color=cyan,author=Pedro]{Describe research design}
\todo[inline,color=cyan,author=Pedro]{Improve RQ descriptions}

\begin{description}
\item[RQ1:] \textit{Which API is easier to learn and use according to the student perceptions?}
\item[RQ2:] \textit{Which API is better to improve the performance according to the student perceptions?}
\end{description}

\subsection{Students' Background}

\todo[inline,color=cyan,author=Pedro]{Describe students' background}

\begin{figure}[htpb]
    \centering
    \includegraphics[width=0.85\columnwidth]{background_questions}
    \caption{Student mean knowledge \textit{before} the assignment}
    \label{fig:background}
\end{figure}

\begin{table}[htpb]
    \centering
    \begin{tabular}{@{}p{0.9\columnwidth}p{0.08\columnwidth}@{}}
        \toprule
        \multicolumn{1}{c}{\scriptsize{Have you$\dots$}} & \textnumero \\ \midrule
        \scriptsize{Had contact with parallel and distributed concepts before?} & $(1)$ \\
        \scriptsize{Had contat with the APIs before?} & $(2)$ \\
        \scriptsize{Learned from the OpenMP class?} & $(3)$ \\
        \scriptsize{Learned from the Pthreads class?} & $(4)$ \\ \bottomrule
    \end{tabular}
    \caption{Questions for Figure \ref{fig:classes}}
    \label{tab:classes}
\end{table}

\begin{figure}[htpb]
    \centering
    \includegraphics[width=0.85\columnwidth]{classes_questions}
    \caption{Student API knowledge \textit{before} the assignment and their relation to classes}
    \label{fig:classes}
\end{figure}

>>>>

\todo[inline,color=cyan,author=Pedro]{Describe research design}

\subsection{The Concurrent, Parallel and Distributed Computing Course}

\todo[inline,color=cyan,author=Pedro]{Describe the course}

\subsection{Questionnaire}

\todo[inline,color=cyan,author=Pedro]{Describe the questionnaire}
\todo[inline,color=cyan,author=Pedro]{Add reference for questionnaires from Herez}

\subsection{Research Questions}

\todo[inline,color=cyan,author=Pedro]{Improve RQ discussion and descriptions}

According to the students' perception, which API:

\begin{description}
    \item[RQ1:] \textit{Was easier to learn and use?}
    \item[RQ2:] \textit{Improved performance the most?}
\end{description}

After evaluation of the students' performance improvements
in the assignment:

\begin{description}
    \item[RQ3:] \textit{Did the students' performance improvements match their
        perception?}
\end{description}

### 2017-08-01 sections/05-responses.tex
Improve responses section.

\section{Students' Responses}
\label{sec:responses}

This section presents the responses given by the students to our
questionnaire. The questions are listed in Tables \ref{tab:likert}
and \ref{tab:comparisons}.

\todo[inline,color=cyan,author=Pedro]{Add reference for questionnaires from Herez}
\todo[inline,color=cyan,author=Pedro]{Improve figure colorscheme}

The first set of questions used a Likert-type Scale to measure the relationship
of the students with the two APIs in the context of the assignment. The Likert
Scale~\cite{likert1932technique} enables respondents to express their agreement
or disagreement with specific statements. Respondents choose between five
options: \textit{Strongly Agree} (SA), \textit{Agree} (A), \textit{Neutral}
(N), \textit{Disagree} (D) and \textit{Strongly Disagree}.

>>>>

\section{Students' Background \& Responses}
\label{sec:responses}

In this section we present the background of our students in the technologies
relevant to the course and to the assignment. Then we present the responses
given by the students to our questionnaire. The questions are listed in Tables
\ref{tab:likert} and \ref{tab:comparisons}.

\todo[inline,color=cyan,author=Pedro]{Improve figure colorscheme}

\subsection{Background}

The course of Concurrent, Parallel and Distributed Computing is offered to
students of both the graduate and undergraduate programs in Computer Science at
the University of São Paulo. To gauge the proficiency and knowledge of our
students we elaborated a five-point scale from \textit{Do not Know} $(1)$ to
\textit{Expert or Proficient} $(5)$.  The students were asked to express their
proficiency level on this scale, and their responses are shown in Figure
\ref{fig:background}.

\begin{figure}[htpb]
    \centering
    \includegraphics[width=0.85\columnwidth]{background_questions}
    \caption{Student knowledge \textit{before} the assignment}
    \label{fig:background}
\end{figure}

It is common for Computer Science introductory and advanced courses in the
graduate and undergraduate programs of the University of São Paulo to be given
in programming languages such as C, Java and Python. Figure
\ref{fig:background} shows that students reported higher proficiency in those
languages. The C language was the one with which students reported most
proficiency.

The students were asked to evaluate their proficiency in \textit{OpenMP} and
\textit{Pthreads}, \textit{before} completing the assignment, with two sets of
questions. The first set used the same five-point scale as the other background
questions and is also shown in Figure \ref{fig:background}. The students
reported significantly larger proficiency in \textit{Pthreads} than
\textit{OpenMP}, and the majority of the students reported they did not know
\textit{OpenMP} before the course.

\begin{table}[htpb]
    \centering
    \begin{tabular}{@{}p{0.9\columnwidth}p{0.08\columnwidth}@{}}
        \toprule
        \multicolumn{1}{c}{\scriptsize{Have you$\dots$}} & \textnumero \\ \midrule
        \scriptsize{Had contact with parallel and distributed concepts before?} & $(1)$ \\
        \scriptsize{Had contat with the APIs before?} & $(2)$ \\
        \scriptsize{Learned from the OpenMP class?} & $(3)$ \\
        \scriptsize{Learned from the Pthreads class?} & $(4)$ \\ \bottomrule
    \end{tabular}
    \caption{Questions for Figure \ref{fig:classes}}
    \label{tab:classes}
\end{table}

The second set of questions regarding student proficiency with the APIs
comprised the 4 \textit{Yes or No} questions listed in Table \ref{tab:classes}.
This set of questions also asked the students to evaluate whether the classes
on each subject helped their understanding and learning of the technologies.
The student responses for this set are shown in Figure \ref{fig:classes}.  The
majority of the students have had previous contact with the concepts related to
concurrent, parallel and distributed programming, but the majority also have
not had contact or experience with \textit{OpenMP} or \textit{Pthreads}.  Most
students reported that the classes on each API helped their learning.

\begin{figure}[htpb]
    \centering
    \includegraphics[width=0.85\columnwidth]{classes_questions}
    \caption{Student API knowledge \textit{before} the assignment and their relation to classes}
    \label{fig:classes}
\end{figure}

\subsection{Responses}

The students were asked to answer 3 sets of questions regarding their
experience with \textit{OpenMP} and \textit{Pthreads}, \textit{after}
completing the assignment.
The first set of questions used a Likert-type Scale to measure the relationship
of the students with the two APIs in the context of the assignment. The Likert
Scale~\cite{likert1932technique} enables respondents to express their agreement
or disagreement with specific statements. Respondents choose between five
options: \textit{Strongly Agree} (SA), \textit{Agree} (A), \textit{Neutral}
(N), \textit{Disagree} (D) and \textit{Strongly Disagree}.
